\documentclass[12pt,letterpaper,boxed,cm]{hmcpset}

\usepackage[margin=1in]{geometry}
\usepackage{mathtools}
\usepackage{mathrsfs}
\usepackage{graphicx}


\name{A GIRL HAS NO NAME}
\class{Computer Science 81}
\assignment{Homework 2}
\duedate{1/31/17}

\newcommand{\pn}[1]{\left( #1 \right)}
\newcommand{\abs}[1]{\left| #1 \right|}
\newcommand{\bk}[1]{\left[ #1 \right]}
\newcommand{\ra}[0]{\rightarrow}
\newcommand{\cp}[0]{~\vdash~}

\begin{document}
\problemlist{1, 2, 3, 4, 5}

\begin{problem}[1]
    [20 points] Give box-style proofs for each of the following:
    \begin{enumerate}
        \item [A.] [4 points] $F \ra G \cp \pn{\pn{F \land G} \ra F} \land \pn{F \ra \pn{F \land G}}.$
        \item [B.] [4 points] $\pn{F \ra G} \land \pn{F \ra H} \cp F \ra \pn{G \land H}.$
        \item [C.] [4 points] $F \lor G, ~\neg F \cp G.$
        \item [D.] [4 points] $\neg\pn{F \ra G} \cp G \ra F.$
        \item [E.] [4 points] $\neg\pn{F \land G} \cp \neg F \lor \neg G.$
    \end{enumerate}
\end{problem}

\begin{solution}
    \vfill
\end{solution}
\newpage

\begin{problem}[2]
    [6 points] Pierce's Law is the formula $\pn{\pn{F \ra G} \ra F} \ra F.$
    \begin{enumerate}
        \item [A.] [2 points] By using truth tables, show that Pierce's law is a tautology (i.e., that  $~\vDash \pn{\pn{F \ra G} \ra F} \ra F$), or equivalently that this formula is true under every valuation.
        \item [B.] [4 points] Prove Pierce's Law using natural deduction.
        \begin{itemize}
            \item [] Hint: As mentioned in class, \textit{constructive} logic rejects the use of “indirect” arguments. Consequently, in constructive logic one cannot use the excluded middle axiom $\pn{\vdash~\phi\lor\neg\phi}$, double-negation elimination rule $\pn{\neg\neg\phi \cp\phi} $, or the proof by contradiction rule (if $\neg\phi\cp\bot$, then$\cp\phi$). In constructive logic, Pierce's Law is not provable!  Therefore, your proof has to involve at least one use of at least one of these three classical rules.
        \end{itemize}
    \end{enumerate}
    
\end{problem}

\begin{solution}
    \vfill
\end{solution}
\newpage

\begin{problem}[3]
    [24 points] \textbf{Proof or No Proof?} For each of the following, either show there is no proof (by finding a valuation/model that makes the assumptions true and the conclusion false), or provide a natural deduction (box) proof.
    \begin{enumerate}
        \item [A.] [4 points] $\neg F\lor\pn{G\ra F}\cp\neg F\land G.$
        \item [B.] [4 points] $F \ra G,~\neg F\ra G \cp G.$
        \item [C.] [4 points] $F \ra G \ra H \cp G \ra F \ra H.$
        \item [D.] [4 points] $F \ra G \ra H \cp F \ra H \ra G.$
        \item [E.] [4 points] $\pn{F \ra G} \ra H \cp F \ra G \ra H.$
        \item [F.] [4 points] $F \ra G,~C \ra D \cp F \lor C \ra G \land D.$
    \end{enumerate}
\end{problem}

\begin{solution}
    \vfill
\end{solution}
\newpage

\begin{problem}[4]
    [9 points] \textbf{OS Wars:} Consider the following argument:
    \begin{itemize}
        \item Premise 1: If my computer runs Windows, then Microsoft got my money.
        \item Premise 2: If my computer runs Mac OS X, then Apple got my money.
        \item Conclusion: At least one of the following statements is true:
        \begin{itemize}
            \item If my computer runs Mac OS X, then Microsoft got my money.
            \item If my computer runs Windows, then Apple got my money.
        \end{itemize}
    \end{itemize}
    \begin{enumerate}
        \item [A.] Define propositions (e.g., $F$, $G$, etc.) to stand for relevant statements in this proof, and show how to symbolically represent the premises and conclusion using these propositions.
        \item [B.] Either provide a formal proof that this argument holds, or find a valuation where the assumptions are true but the conclusion is false.
    \end{enumerate}
\end{problem}

\begin{solution}
    \vfill
\end{solution}
\newpage

\begin{problem}[5]
    [1 easy point] Please wait until you're done with the rest of the assignment to answer this quick survey:
    \begin{enumerate}
        \item [A.] How long (in hours) did you spend working on this assignment?
        \item [B.] What was the most interesting thing you learned while answering these problems? (We're sure there was \textit{something} you learned.)
    \end{enumerate}
\end{problem}

\begin{solution}
    \vfill
\end{solution}
\newpage

\end{document}